---
title: "Download functions for Meteogalicia"
author: "Dominic Royé"
date: "20/02/2019"
output: rmarkdown::github_document
---

At this moment two functions are available to download data of the Galician Weather Service. 

1) Lightning strike data.

2) Daily meteorological variables.

## Meteorological data

```{r,warning=FALSE,message=FALSE}
#install the libraries if necessary
if(!require("tidyverse")) install.packages("tidyverse")
if(!require("XML")) install.packages("XML")
if(!require("jsonlite")) install.packages("jsonlite")
if(!require("sf")) install.packages("sf")
if(!require("rnaturalearth")) install.packages("rnaturalearth")


#packages for functions
library(tidyverse) #+plotting
library(XML)
library(jsonlite)

#packages for plotting
library(sf)
library(rnaturalearth)

#functions
source("meteo_daily_download.R") #meteo download
source("lightning_download.R") #lightning download

#station details
load("altabaja.RData")
load("idestaciones.RData")
load("idparametros.RData")

```

The following tables include important details about station ids, longitude, latitude, operative stations, available variables, etc. 

```{r}
#stations id + latlon + place
str(idestaciones)

#information about which stations are operative (FechAlta: from, FechaBaja: to)
str(altabaja)

#id_names for all available variables
str(idparametros)

```

### Example 

Lugo-Campus with maximum temperature (1.5m) in January 2019.

```{r,warning=FALSE,message=FALSE,out.width="100%"}
#downloading data
data <- import_meteo_daily(10053,         #station id
                           "TA_MAX_1.5m", #variable id
                           "01/01/2019",  #date from
                           "31/01/2019")  #date to 

#data structure
str(data)

#plotting time serie
ggplot(data,
       aes(date,valor))+
          geom_line()
```

It is possible to query several stations or variables at the same time. However, the limit is set at 3 months for a single station and variable.

```{r,warning=FALSE,message=FALSE,out.width="100%"}
#downloading data
data <- import_meteo_daily(c(10053,14000), #station id
                           "TA_MAX_1.5m",  #variable id
                           "01/01/2019",   #date from
                           "31/01/2019")   #date to 

#data structure
str(data)

#plotting time serie
ggplot(data,
       aes(date,valor))+
          geom_line()+
        facet_grid(estacion~.)


```


## Lightning strikes

```{r,warning=FALSE,message=FALSE,out.width="100%"}
#download lightning strikes
data <- lightning_download("01/01/2019","31/01/2019")

#data structure
str(data)

#lonlat to numeric
data <- mutate(data,lon=as.numeric(as.character(lon)),
               lat=as.numeric(as.character(lat)))

data_sf <- st_as_sf(data,coords=c("lon","lat"))%>%
               st_set_crs(4326)

map_bg <- ne_countries(scale=10,returnclass = "sf")

#plotting
     ggplot()+
        geom_sf(data=map_bg,
                fill="black",
                colour="white",
                size=0.2)+
         geom_sf(data=data_sf,
                 colour="#cb181d",
                 alpha=0.5)+
           coord_sf(xlim=c(-16,5),
                    ylim=c(35,49))+
            theme_minimal()


```
